
# Cleaning The Data 
Now it is time to clean up the dataframe `NYT_raw`. 

First thing first, we start by loading the packages `tidyverse` [@tidyverse2019] and`DT` [@R-DT].
```{r}
pacman::p_load(tidyverse, DT)
```

Here we are going to make the cleaned dataset `NYT_clean` also refered to as `df_NYT` from the dataset `NYT_raw`. It mostly follows the syntax from the package `tidyverse`. 
```{r cleaning the data, eval = F}
#selecting columns to keep
coloumns_to_select <- c("response.docs.headline.main",
"response.docs.web_url", "response.docs.pub_date")

#creating the new dataframe
df_NYT <- NYT_raw %>% 
  #selecting the defined columns
  select(coloumns_to_select) %>% 
  
  #renaming columns to more humane names
  rename(
    "headline" = "response.docs.headline.main",
    "url" = "response.docs.web_url",
    "date" = "response.docs.pub_date",
  ) %>%
  
  #formatting columns to the correct class
  mutate(
    date = as.Date(date)
  ) %>% 
  
  #arranging by date so that the articles are in chronological order
  arrange(by=date)


#saving to a csv
write_csv(df_NYT, "data/new_york_times/data_additional/NYT_clean_cp2.csv")
```

Now lets inspect the cleaned dataframe. to see if it looks allright. To do this we use the function `datatable` from the package `DT`. The cleaned dataset contains the following columns:

* `headline` which is the title/headline of the article
* `url` which is a url leading to the the article on the NYT webpage
* `date` which is the publication date

```{r, message = F, warning = F}
#reading the csv
df_NYT <- read_csv("data/new_york_times/data_additional/NYT_clean_cp2.csv")

#making a nice dataframe that we can browse. Note that we remove abstract because there is too much text in it to show in a nice way.
font.size <- "8pt"
   DT::datatable(
     df_NYT,
     rownames = FALSE,
     filter = "top", 
     options = list(
       initComplete = htmlwidgets::JS(
          "function(settings, json) {",
          paste0("$(this.api().table().container()).css({'font-size': '", font.size, "'});"),
          "}"),
       pagelength = 3,
       scrollX=T,
       autoWidth = TRUE
       ) 
     )
```












