
# Cleaning the Data

<style>
div.blue { background-color:#ffd966; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Note: This chapter is very similar to chapter 5. Here all the steps are just adjusted to the fit the dataset from The Guardian-
</div>

First thing first, we start by loading the packages `tidyverse` [@tidyverse2019] and`DT` [@R-DT]
```{r}
pacman::p_load(tidyverse, DT)
```

Now it is time to clean up the dataframe `guardian_raw` that we made in chapter \@ref

In this chunk we are going to make the cleaned dataset `guardian_clean` also refered to as `df_guardian` from the dataset `guardian_raw`. It mostly follows the syntax from the package `tidyverse`. 
```{r cleaning the data2, eval = F}
#selecting columns to keep
coloumns_to_select <- c("response.results.webTitle", "response.results.webUrl", "response.results.webPublicationDate")

#creating the new dataframe
df_guardian <- guardian_raw %>% 
  
  #selecting the defined columns
  select(coloumns_to_select) %>% 
  
  #renaming columns to more humane names
  rename(
    "headline" = "response.results.webTitle",
    "url" = "response.results.webUrl",
    "date" = "response.results.webPublicationDate",
  ) %>%
  
  #formatting columns to the correct class
  mutate(
    date = as.Date(date),
  ) %>% 
  
  #arranging by date so that the articles are in chronological order
  arrange(by=date)

#saving to a csv
write_csv(df_guardian, "data/guardian/data_additional/guardian_clean_cp2.csv")
```

Now lets inspect the cleaned dataframe. to see if it looks allright. To do this we use the function `datatable` from the package `DT`.

```{r, message = F, warning = F}
#reading the csv
df_guardian <- read_csv("data/guardian/data_additional/guardian_clean_cp2.csv")

#making a nice dataframe that we can browse.
font.size <- "8pt"

   DT::datatable(
     df_guardian,
     rownames = FALSE,
     filter = "top", 
     options = list(
       initComplete = htmlwidgets::JS(
          "function(settings, json) {",
          paste0("$(this.api().table().container()).css({'font-size': '", font.size, "'});"),
          "}"),
       pagelength = 3,
       scrollX=T,
       autoWidth = TRUE
       ) 
     )
```
The cleaned dataset contains the following columns:

* `headline` which is the title/headline of the article
* `url` which is a url leading to the the article on The Guardian webpage
* `date` which is the publication date

All Guchi. 











