# Persons Vizualisation

In this chapter I will make visualizations of key persons to see in which periods of time that these key persons were relevant for the Taliban-conflict. I also compare the two newspapers to see if they focus more or less on these key persons. 

The code in this chapter is mainly for making plots. The commenting will be sparse as it would be too tidious to comment everything. 


First thing first, we start by loading the package `tidyverse` [@tidyverse2019], `patchwork` [@R-patchwork] and `wesanderson` [@R-wesanderson].
```{r}
pacman::p_load(tidyverse, patchwork, wesanderson)
```

Then we load the data containing persons.
```{r, message=F}
person_nyt <- read_csv("data/new_york_times/data_NER/person.csv")
person_guardian <- read_csv("data/guardian/data_NER/person.csv")
```

We do a quick inspection of the most important persons.
```{r, results = F}
z <- person_nyt %>%
  group_by(person) %>% 
  summarize(n = sum(count)) %>% 
  arrange(desc(n))

g <- person_guardian %>%
  group_by(person) %>% 
  summarize(n = sum(count)) %>% 
  arrange(desc(n))
```

The following key persons will be plotted:

* **United States presidents**
   * *George W. Bush*
   * *Barack Obama*
   * *Donald Trump*
* **British Prime Minister**
   * *Tony Blair*
   * *Gordon Brown*
   * *David Cameron*
   * *Boris Johnson*
* **Afghanistan presidents**
   * *Hamid Karzai*
   * *Ashraf Ghani*
* **Other Persons**
   * *Osama bin Laden*
   * *Saddan Hussein*
   

<style>
div.red { background-color:#e06666; border-radius: 5px; padding: 20px;}
</style>
<div class = "red">
Note that Theresa May, British prime minister between 2016-06-13 and 2019-07-24, was not picked up by spaCy and she will therefore not be plotted. 
</div>


## Standardize to hits pr. article and grouping by month

In order to compare the two newspapers we need to standardize the count according to the number of articles pr. month. We will make a new variable called `penalized_count` for both `person_nyt` and `person_guardian`.  

We start by adding a new column `month` to `person_nyt` and `person_guardian`.
```{r}
person_nyt <- person_nyt %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))

person_guardian <- person_guardian %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))
```

Then we load the dataset containing articles for both newspapers as `df_nyt` and `df_guardian`. 
```{r, message = F}
df1 <- read_csv("data/new_york_times/NYT_clean_1.csv")
df2 <- read_csv("data/new_york_times/NYT_clean_2.csv")
df3 <- read_csv("data/new_york_times/NYT_clean_3.csv")

df <- rbind(df1, df2, df3)
df_nyt <- as_tibble(df)

df_guardian <- read_csv("data/guardian/guardian_clean.csv")
```

Here we add a new column to each dataset called `month` .
```{r}
df_nyt <- df_nyt %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))

df_guardian <- df_guardian %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))
```

Then we add a new column `sum_articles_month` which is the sum of articles for a given month. 
```{r}
df_nyt <- df_nyt %>% group_by(month) %>% 
  tally() %>% 
  rename(sum_articles_month = n)

df_guardian <- df_guardian %>% group_by(month) %>% 
  tally() %>% 
  rename(sum_articles_month = n)
```

For the person datasets, we group by person and month and summarize the count. So we get a score for each person for each month.  
```{r, message = F, warning = F}
person_nyt <- person_nyt %>% 
  group_by(person, month) %>% 
  summarise(count = sum(count))

person_guardian <- person_guardian %>% 
  group_by(person, month) %>% 
  summarise(count = sum(count))
```

Then we merge `df_nyt` with `person_nyt` so that we now have a column in `person_nyt` called `sum_articles_month` which indicates the total number of articles published in the given month. I do the same for `df_guardian` and `person_guardian`. 
```{r}
person_nyt <- left_join(person_nyt, df_nyt, by = "month")
person_guardian <- left_join(person_guardian, df_guardian, by = "month")
```

Then we make the columns `penalized_count` and `penalized_count_round` which indicates the number of hits for a given person pr. month pr. total articles in that month.We also make a column `newspaper` to indicate the newspaper. Some months have very few articles, so we also filter out out these, because they can make unreliable values of `penalized_count`.  
```{r}
person_nyt <- person_nyt %>% 
  mutate(
    penalized_count = count/sum_articles_month,
    penalized_count_round = round(penalized_count, 1),
    newspaper = "New York Times"
  ) %>% 
  filter(sum_articles_month > 15)

person_guardian <- person_guardian %>% 
  mutate(
    penalized_count = count/sum_articles_month,
    penalized_count_round = round(penalized_count, 1),
    newspaper = "The Guardian"
  ) %>% 
  filter(sum_articles_month > 15)
```

Then i bind the datasets together so we can plot them in the same plot and select the columns needed.
```{r}
df_person <- rbind(person_nyt, person_guardian)
df_person <- as_tibble(df_person) 
```

Now we have a dataset that is ready for plots. 

## Color Palettes

Here I quickly define some color palettes with colors that i like. These palettes will be used for different plots. The same colors will be used consistently throughout the notebook.
```{r}
color_palette_newspaper <- c(wes_palette("Chevalier1")[1], wes_palette("Darjeeling2")[2])
color_palette_afghan_pres <- c(wes_palette("Moonrise1")[2], wes_palette("Moonrise3")[3])
color_palette_terrorists <- c(wes_palette("Royal1")[4], wes_palette("Royal2")[5])
```


## United States Presidents

### Summary plot
```{r}
summary_nyt <- df_person %>% 
  filter(newspaper == "New York Times") %>% 
  filter(person %in% c("Bush", "Obama", "Trump")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/21109) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[1]) + 
  coord_flip() + 
  labs(x= "President", y="Hits per Article", title = "New York Times") + 
  theme_minimal() + 
  ylim(0,1)

summary_guardian <- df_person %>% 
  filter(newspaper == "The Guardian") %>% 
  filter(person %in% c("Bush", "Obama", "Trump")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/9609) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[2]) + 
  coord_flip() + 
  labs(x= "President", y="Hits per Article", title = "The Guardian") + 
  theme_minimal() + 
  ylim(0,1)

u <- summary_nyt / summary_guardian +
  plot_annotation(title = "U.S. Presidents",
                  theme = theme(plot.title = element_text(size=18)))


```


### George W. Bush
```{r, message = F}
bush <- df_person %>% 
  filter(person == "Bush") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "George W. Bush") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-01-20"), as.Date("2009-01-20"))),
                          linetype = c("solid", "longdash", "solid", "longdash"))
  

bush
```

### Barack Obama

```{r, message = F}
obama <- df_person %>% 
  filter(person == "Obama") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Barack Obama") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2004-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2009-01-20"), as.Date("2017-01-20"))),
                          linetype = c("solid", "longdash", "solid", "longdash"))

obama
```

### Donald Trump

```{r, message = F}
trump <- df_person %>% 
  filter(person == "Trump") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Donald Trump") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2004-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2017-01-20"), as.Date("2021-01-20"))),
                          linetype = c("solid", "longdash", "solid", "longdash"))

trump
```

### Combined Plots

#### New York Times

United states president plot for New York Times
```{r, message = F, fig.dim=c(10,5)}
bush <- df_person %>% 
  filter(person == "Bush") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "George W. Bush"
  ) 

obama <- df_person %>% 
  filter(person == "Obama") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Barack Obama"
  )

trump <- df_person %>% 
  filter(person == "Trump") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Donald Trump"
  )

pres <- rbind(bush, obama, trump)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, c("George W. Bush", "Barack Obama")))

pres_nyt <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "U.S Presidents Mentioned in New York Times") +
  scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "FantasticFox1"), name = "")

pres_nyt
```

#### The Guardian

United States president plot for the Guardian.
```{r, message = F, fig.dim=c(10,5)}
bush <- df_person %>% 
  filter(person == "Bush") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "George W. Bush"
  ) 

obama <- df_person %>% 
  filter(person == "Obama") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Barack Obama"
  )

trump <- df_person %>% 
  filter(person == "Trump") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Donald Trump"
  )

pres <- rbind(bush, obama, trump)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, c("George W. Bush", "Barack Obama")))

pres_guardian <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "U.S Presidents Mentioned in The Guardian") +
  scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "FantasticFox1"), name = "") 

pres_guardian
```

#### Both Newspapers Without Facet Wrap
```{r, message = F, warning = F}
pres_nyt_y <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,7) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-01-20"), as.Date("2009-01-20"), as.Date("2017-01-20"))),
                          linetype = c("solid", "solid", "solid"),
                          color = c(wes_palette("FantasticFox1")[1], wes_palette("FantasticFox1")[2], wes_palette("FantasticFox1")[3])) 


pres_guardian_y <- pres_guardian +
  theme(legend.position="bottom") +
  ylim(-0.5,7) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-01-20"), as.Date("2009-01-20"), as.Date("2017-01-20"))),
                          linetype = c("solid", "solid", "solid"),
                          color = c(wes_palette("FantasticFox1")[1], wes_palette("FantasticFox1")[2], wes_palette("FantasticFox1")[3])) 


#combining into single plot
pres_both_newspapers <- pres_nyt_y / pres_guardian_y

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers
```

#### Both Newspapers With Facet Wrap
```{r, message = F, warning = F}
pres_nyt <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,3) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "New York Times")
  
pres_guardian <- pres_guardian + 
  theme(legend.position="none") + 
  ylim(-0.5,3) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "5 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "The Guardian")


#combining into single plot
pres_both_newspapers <- pres_nyt / pres_guardian

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers +
  plot_annotation(title = "Timeline of U.S. Presidents",
                  theme = theme(plot.title = element_text(size=18)))
```

## British Prime Ministers

<style>
div.red { background-color:#e06666; border-radius: 5px; padding: 20px;}
</style>
<div class = "red">
Note that Theresa May, British prime minister between 2016-06-13 and 2019-07-24, was not picked up by spaCy and she will therefore not be plotted. 
</div>

### Summary plot
```{r}
summary_nyt <- df_person %>% 
  filter(newspaper == "New York Times") %>% 
  filter(person %in% c("Blair", "Brown", "Cameron", "Boris")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/21109) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[1]) + 
  coord_flip() + 
  labs(x= "Prime Minister", y="Hits per Article", title = "New York Times") + 
  theme_minimal() + 
  ylim(0,0.35)

summary_guardian <- df_person %>% 
  filter(newspaper == "The Guardian") %>% 
  filter(person %in% c("Blair", "Brown", "Cameron", "Boris")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/9609) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[2]) + 
  coord_flip() + 
  labs(x= "Prime Minister", y="Hits per Article", title = "The Guardian") + 
  theme_minimal() + 
  ylim(0,0.35)

b <- summary_nyt / summary_guardian +
  plot_annotation(title = "British Prime Ministers",
                  theme = theme(plot.title = element_text(size=18)))

b
```

### Tony Blair

```{r, message = F, warning = F}
blair <- df_person %>% 
  filter(person == "Blair") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Tony Blair") +
  facet_wrap(~newspaper, nrow = 2) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("1997-05-02"), as.Date("2007-06-27"))),
                          linetype = c("solid", "longdash", "solid", "longdash")) +
  scale_x_date(limits = as.Date(c("1997-01-01", "2021-11-01"))) 

blair
```
### Gordon Brown

```{r, message = F, warning = F}
brown <- df_person %>% 
  filter(person == "Brown") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Gordon Brown") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(date_breaks = "3 year") +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2007-06-27"), as.Date("2010-05-11"))),
                          linetype = c("solid", "longdash", "solid", "longdash")) +
  scale_x_date(limits = as.Date(c("1997-01-01", "2021-11-01"))) 

brown
```

### David Cameron

```{r, message = F, warning = F}
cameron <- df_person %>% 
  filter(person == "Cameron") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "David Cameron") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(date_breaks = "3 year") +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2010-05-11"), as.Date("2016-06-13"))),
                          linetype = c("solid", "longdash", "solid", "longdash")) +
  scale_x_date(limits = as.Date(c("2000-01-01", "2021-11-01"))) 

cameron
```

### Boris Johnson

```{r, message = F, warning = F}
boris <- df_person %>% 
  filter(person == "Boris") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Boris Johnson") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(date_breaks = "3 year") +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") +
  geom_vline(xintercept = as.numeric(c(as.Date("2019-06-24"))),
                          linetype = c("solid", "solid")) +
  scale_x_date(limits = as.Date(c("1997-01-01", "2021-11-01"))) 

boris
```
### Combined Plots

#### New York Times

British prime minister plot for New York Times.
```{r, message = F, fig.dim=c(10,5)}
blair <- df_person %>% 
  filter(person == "Blair") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Tony Blair"
  ) 

brown <- df_person %>% 
  filter(person == "Brown") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Gordon Brown"
  )

cameron <- df_person %>% 
  filter(person == "Cameron") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "David Cameron"
  )

boris <- df_person %>% 
  filter(person == "Boris") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Boris Johnson"
  )

pres <- rbind(blair, brown, cameron, boris)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, c("Tony Blair", "Gordon Brown", "David Cameron")))

pres_nyt <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.7) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "British Prime Ministers Mentioned in New York Times") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "Darjeeling1"), name = "")

pres_nyt
```

#### The Guardian

British prime minister plot for the Guardian.
```{r, message = F, fig.dim=c(10,5)}
blair <- df_person %>% 
  filter(person == "Blair") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Tony Blair"
  ) 

brown <- df_person %>% 
  filter(person == "Brown") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Gordon Brown"
  )

cameron <- df_person %>% 
  filter(person == "Cameron") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "David Cameron"
  )

boris <- df_person %>% 
  filter(person == "Boris") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Boris Johnson"
  )

pres <- rbind(blair, brown, cameron, boris)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, c("Tony Blair", "Gordon Brown", "David Cameron")))

pres_guardian <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.7) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "British Prime Ministers Mentioned in The Guardian") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "Darjeeling1"), name = "")

pres_guardian
```

#### Both Newspapers Without Facet Wrap
```{r, message = F, warning = F}
pres_nyt_y <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,4) +
  #scale_x_date(limits = as.Date(c("1997-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("1997-05-02"), as.Date("2007-06-27"), as.Date("2010-05-11"), as.Date("2019-06-24"))),
                          linetype = c("solid", "solid", "solid", "solid"),
                          color = c(wes_palette("Darjeeling1")[1], wes_palette("Darjeeling1")[2],
                                    wes_palette("Darjeeling1")[3], wes_palette("Darjeeling1")[4])) 


pres_guardian_y <- pres_guardian +
  theme(legend.position="bottom") +
  ylim(-0.5,4) +
  #scale_x_date(limits = as.Date(c("1997-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("1997-05-02"), as.Date("2007-06-27"), as.Date("2010-05-11"), as.Date("2019-06-24"))),
                          linetype = c("solid", "solid", "solid", "solid"),
                          color = c(wes_palette("Darjeeling1")[1], wes_palette("Darjeeling1")[2],
                                    wes_palette("Darjeeling1")[3], wes_palette("Darjeeling1")[4]))


#combining into single plot
pres_both_newspapers <- pres_nyt_y / pres_guardian_y

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers
```

#### Both Newspapers With Facet Wrap
```{r, message = F, warning = F}
pres_nyt <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,4) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "6 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "New York Times")
  
pres_guardian <- pres_guardian + 
  theme(legend.position="none") + 
  ylim(-0.5,4) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "6 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "The Guardian")


#combining into single plot
pres_both_newspapers <- pres_nyt / pres_guardian

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers +
  plot_annotation(title = "Timeline of British Prime Ministers",
                  theme = theme(plot.title = element_text(size=18)))
```

## Afghanistan Presidents

### Summary plot
```{r}
summary_nyt <- df_person %>% 
  filter(newspaper == "New York Times") %>% 
  filter(person %in% c("Karzai", "Ghani")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/21109) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[1]) + 
  coord_flip() + 
  labs(x= "President", y="Hits per Article", title = "New York Times") + 
  theme_minimal() + 
  ylim(0,0.6)

summary_guardian <- df_person %>% 
  filter(newspaper == "The Guardian") %>% 
  filter(person %in% c("Karzai", "Ghani")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/9609) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[2]) + 
  coord_flip() + 
  labs(x= "President", y="Hits per Article", title = "The Guardian") + 
  theme_minimal() + 
  ylim(0,0.6)

a <- summary_nyt / summary_guardian +
  plot_annotation(title = "Afghanistan Presidents",
                  theme = theme(plot.title = element_text(size=18)))

a
```

### Hamid Karzai

```{r, message = F}
karzai <- df_person %>% 
  filter(person == "Karzai") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Hamid Karzai") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2000-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")

karzai
```

### Ashraf Ghani

```{r, message = F}
ghani <- df_person %>% 
  filter(person == "Ghani") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Ashraf Ghani") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")

ghani
```

### Combined Plots

#### New York Times

Afghanistan president plot for New York Times.
```{r, message = F, fig.dim=c(10,5)}
karzai <- df_person %>% 
  filter(person == "Karzai") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Hamid Karzai"
  ) 

ghani <- df_person %>% 
  filter(person == "Ghani") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    president = "Ashraf Ghani"
  )


pres <- rbind(karzai, ghani)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, "Hamid Karzai"))

pres_nyt <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Afghan Presidents Mentioned in New York Times") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_afghan_pres, name = "")

pres_nyt
```

#### The Guardian

Afghanistan president plot for the Guardian.
```{r, message = F, fig.dim=c(10,5)}
karzai <- df_person %>% 
  filter(person == "Karzai") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Hamid Karzai"
  ) 

ghani <- df_person %>% 
  filter(person == "Ghani") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    president = "Ashraf Ghani"
  )

pres <- rbind(karzai, ghani)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(president))

pres <- pres %>% 
  mutate(president = fct_relevel(president, "Hamid Karzai"))

pres_guardian <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = president) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Afghan Presidents Mentioned in The Guardian") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_afghan_pres, name = "")

pres_guardian
```

#### Both Newspapers Without Facet Wrap
```{r, message = F}
pres_nyt_t <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,3.5) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-12-22"), as.Date("2014-09-29"))),
                          linetype = c("solid", "solid"),
                          color = c(color_palette_afghan_pres[1], color_palette_afghan_pres[2]))


pres_guardian_t <- pres_guardian +
  theme(legend.position="bottom") +
  ylim(-0.5,3.5) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-12-22"), as.Date("2014-09-29"))),
                          linetype = c("solid", "solid"),
                          color = c(color_palette_afghan_pres[1], color_palette_afghan_pres[2]))


#combining into single plot
pres_both_newspapers <- pres_nyt_t / pres_guardian_t

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers
```

#### Both Newspapers With Facet Wrap
```{r}
#removing legends
pres_nyt <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,3) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "3 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "New York Times")
  
pres_guardian <- pres_guardian + 
  theme(legend.position="none") + 
  ylim(-0.5,3) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "3 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "The Guardian")


#combining into single plot
pres_both_newspapers <- pres_nyt / pres_guardian

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers +
  plot_annotation(title = "Timeline of Afghanistan Presidents",
                  theme = theme(plot.title = element_text(size=18)))
```

## Other Persons

### Summary plot
```{r}
summary_nyt <- df_person %>% 
  filter(newspaper == "New York Times") %>% 
  filter(person %in% c("Laden", "Hussein")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/21109) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[1]) + 
  coord_flip() + 
  labs(x= "Terrosist", y="Hits per Article", title = "New York Times") + 
  theme_minimal() + 
  ylim(0,0.6)

summary_guardian <- df_person %>% 
  filter(newspaper == "The Guardian") %>% 
  filter(person %in% c("Laden", "Hussein")) %>% 
  group_by(person) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/9609) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(person, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[2]) + 
  coord_flip() + 
  labs(x= "Terrosist", y="Hits per Article", title = "The Guardian") + 
  theme_minimal() + 
  ylim(0,0.6)

t <- summary_nyt / summary_guardian +
  plot_annotation(title = "Terrorists",
                  theme = theme(plot.title = element_text(size=18)))
t
```

### Osama bin Laden

```{r, message = F}
laden <- df_person %>% 
  filter(person == "Laden") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Osama bin Laden") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") + 
  ylim(0,8)

laden
```

### Saddam Hussein

```{r, message = F}
hussein <- df_person %>% 
  filter(person == "Hussein") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Saddam Hussein") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none") 

hussein
```

### Combined Plots

#### New York Times

Terrorist plot for New York Times.
```{r, message = F, fig.dim=c(10,5)}
laden <- df_person %>% 
  filter(person == "Laden") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    terrorist = "Osama bin Laden"
  ) 

hussein <- df_person %>% 
  filter(person == "Hussein") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    terrorist = "Saddam Hussein"
  )


pres <- rbind(laden, hussein)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(terrorist))


pres_nyt <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = terrorist) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Terrorists Mentioned in New York Times") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_terrorists, name = "")

pres_nyt
```

#### The Guardian

Afghanistan president plot for the Guardian.
```{r, message = F, fig.dim=c(10,5)}
laden <- df_person %>% 
  filter(person == "Laden") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    terrorist = "Osama bin Laden"
  ) 

hussein <- df_person %>% 
  filter(person == "Hussein") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    terrorist = "Saddam Hussein"
  )

pres <- rbind(laden, hussein)
pres <- as_tibble(pres) %>% 
  mutate(president = as.factor(terrorist))


pres_guardian <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = terrorist) + 
  geom_point(size = 0.9) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Terrorists Mentioned in The Guardian") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_terrorists, name = "")

pres_guardian
```

#### Both Newspapers Without Facet Wrap
```{r, message = F}
pres_nyt_t <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,3.5) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-12-22"), as.Date("2014-09-29"))),
                          linetype = c("solid", "solid"),
                          color = c(color_palette_afghan_pres[1], color_palette_afghan_pres[2]))


pres_guardian_t <- pres_guardian +
  theme(legend.position="bottom") +
  ylim(-0.5,3.5) +
  #scale_x_date(limits = as.Date(c("1998-01-01", "2021-11-01"))) +
  geom_vline(xintercept = as.numeric(c(as.Date("2001-12-22"), as.Date("2014-09-29"))),
                          linetype = c("solid", "solid"),
                          color = c(color_palette_afghan_pres[1], color_palette_afghan_pres[2]))


#combining into single plot
pres_both_newspapers <- pres_nyt_t / pres_guardian_t

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers
```

#### Both Newspapers With Facet Wrap
```{r}
#removing legends
pres_nyt <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,3) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "3 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "New York Times")
  
pres_guardian <- pres_guardian + 
  theme(legend.position="none") + 
  ylim(-0.5,3) +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "3 years", date_labels = "%Y") +
  facet_wrap(~president, nrow = 1) + 
  labs(title = "The Guardian")


#combining into single plot
pres_both_newspapers <- pres_nyt / pres_guardian

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers +
  plot_annotation(title = "Timeline of Terrorists",
                  theme = theme(plot.title = element_text(size=18)))
```






