# Organisations Vizualisations

In this Chapter I will make vizualisations of key organisations to see which periods of time that these organizations were relevant for the Taliban-conflict. I also compare the two newspapers to see if they focus more or less on these key organizations 

The code in this chapter is mainly for making plots. The commenting will be sparse as it would be too tedious to comment everything. 

First thing first, we start by loading the package `tidyverse` [@tidyverse2019], `patchwork` [@R-patchwork] and `wesanderson` [@R-wesanderson].

```{r}
pacman::p_load(tidyverse, patchwork, wesanderson)
```

Then we load the dataset containing organisations for both newspapers. 
```{r, message = F}
org_nyt <- read_csv("data/new_york_times/data_NER/org.csv")
org_guardian <- read_csv("data/guardian/data_NER/org.csv")
```

Here we get a quick look at what is interesting to pursue in our plots.
```{r, results = F}
org_nyt %>%
  group_by(ORG) %>% 
  summarize(n = sum(count)) %>% 
  arrange(desc(n)) 

org_guardian %>%
  group_by(ORG) %>% 
  summarize(n = sum(count)) %>% 
  arrange(desc(n)) 
```

The following key organizations will be plotted:

* **Western Alliances**
   * *Nato*
* **Militant Islamic Organizations**
   * *Taliban*
   * *Al Queda*
   * *Hamas*
   * *Hezbollah*


## Standardize names of organisations

The name of organizations appears differently. For example NATO appears under several different names such as *Nato* or *NATO*. We will standardize these names so they all appear as the same name.
```{r}
org_nyt <- org_nyt %>% 
  mutate(
    #NATO
    ORG = ifelse(ORG == "nato", "NATO", ORG),
    ORG = ifelse(ORG == "Nato", "NATO", ORG),
    #Al Qaeda
    ORG = ifelse(ORG == "qaeda", "Qaeda", ORG),
    ORG = ifelse(ORG == "Qaida", "Qaeda", ORG),
    #Taliban
    ORG = ifelse(ORG == "taliban", "Taliban", ORG),
    #Hizbollah
    ORG = ifelse(ORG == "Hizbollah", "Hezbollah", ORG),
  )

org_guardian <- org_guardian %>% 
  mutate(
    #NATO
    ORG = ifelse(ORG == "nato", "NATO", ORG),
    ORG = ifelse(ORG == "Nato", "NATO", ORG),
    #Al Qaeda
    ORG = ifelse(ORG == "qaeda", "Qaeda", ORG),
    ORG = ifelse(ORG == "Qaida", "Qaeda", ORG),
    #Taliban
    ORG = ifelse(ORG == "taliban", "Taliban", ORG),
    #Hizbollah
    ORG = ifelse(ORG == "Hizbollah", "Hezbollah", ORG),
  )
```


## Standardize to hits pr. article and grouping by month

In order to compare the two newspapers we need to standardize the count according to the number of articles. We will make a new variable called `penalized_count` for both `org_nyt` and `org_guardian`.  

We start by adding a new column `year` to `org_nyt` and `org_guardian`.
```{r}
org_nyt <- org_nyt %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))

org_guardian <- org_guardian %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))
```

Then we load the dataset containing articles for both newspapers as `df_nyt` and `df_guardian`. 
```{r, message = F}
df1 <- read_csv("data/new_york_times/NYT_clean_1.csv")
df2 <- read_csv("data/new_york_times/NYT_clean_2.csv")
df3 <- read_csv("data/new_york_times/NYT_clean_3.csv")

df <- rbind(df1, df2, df3)
df_nyt <- as_tibble(df)

df_guardian <- read_csv("data/guardian/guardian_clean.csv")
```

Here we add a new column to each dataset called `month` .
```{r}
df_nyt <- df_nyt %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))

df_guardian <- df_guardian %>% 
  mutate(month = lubridate::floor_date(date, "month"),
         month = as.Date(month))
```

Then we add a new column `sum_articles_month` which is the sum of articles for a given year. 
```{r}
df_nyt <- df_nyt %>% group_by(month) %>% 
  tally() %>% 
  rename(sum_articles_month = n)

df_guardian <- df_guardian %>% group_by(month) %>% 
  tally() %>% 
  rename(sum_articles_month = n)
```

For the organisations datasets,We group  by organisation and year and summarize the count. So we get a score for each country for each month. 
```{r}
org_nyt <- org_nyt %>% 
  group_by(ORG, month) %>% 
  summarise(count = sum(count))

org_guardian <- org_guardian %>% 
  group_by(ORG, month) %>% 
  summarise(count = sum(count))
```

Then we merge `df_nyt` with `org_nyt` so that we now have a column in `org_nyt` called `sum_articles_month` which indicates the total number of articles published in the given month. I do the same for `df_guardian` and `org_guardian`. 
```{r}
org_nyt <- left_join(org_nyt, df_nyt, by = "month")
org_guardian <- left_join(org_guardian, df_guardian, by = "month")
```

Then we make the columns `penalized_count` and `penalized_count_round` which indicates the number of hits for a given person pr. month pr. total articles in that month.We also make a column `newspaper` to indicate the newspaper. Some months have very few articles, so we also filter out out these, because they can make unreliable values of `penalized_count`.  
```{r}
org_nyt <- org_nyt %>% 
  mutate(
    penalized_count = count/sum_articles_month,
    penalized_count_round = round(penalized_count, 1),
    newspaper = "New York Times"
  ) %>% 
  filter(sum_articles_month > 30)

org_guardian <- org_guardian %>% 
  mutate(
    penalized_count = count/sum_articles_month,
    penalized_count_round = round(penalized_count, 1),
    newspaper = "The Guardian"
  ) %>% 
  filter(sum_articles_month > 30)
```

Then i bind the datasets together so we can plot them in the same plot and select the columns needed.
```{r}
df_org <- rbind(org_nyt, org_guardian)
df_org <- as_tibble(df_org) 
```

Now we have a dataset that is ready for plots. 

## Color Palettes

Here I quickly define some color palettes with colors that i like. These palettes will be used for different plots. The same colors will be used consistently throughout the notebook.
```{r}
color_palette_newspaper <- c(wes_palette("Chevalier1")[1], wes_palette("Darjeeling2")[2])
```

## Summary plot
```{r}
summary_nyt <- df_org %>% 
  filter(newspaper == "New York Times") %>% 
  filter(ORG %in% c("NATO", "Taliban", "Qaeda", "Hamas", "Hezbollah")) %>% 
  group_by(ORG) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/21109) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(ORG, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[1]) + 
  coord_flip() + 
  labs(x= "Organisation", y="Hits per Article", title = "New York Times") + 
  theme_minimal() + 
  ylim(0,4.2)

summary_guardian <- df_org %>% 
  filter(newspaper == "The Guardian") %>% 
  filter(ORG %in% c("NATO", "Taliban", "Qaeda", "Hamas", "Hezbollah")) %>%
  group_by(ORG) %>% 
  summarise(count = sum(count)) %>%  
  mutate(hits_pr_article = count/9609) %>% 
  arrange(desc(hits_pr_article)) %>% 
  ggplot() + 
  aes(x=reorder(ORG, hits_pr_article), y=hits_pr_article) +
  geom_bar(stat = "identity", fill = color_palette_newspaper[2]) + 
  coord_flip() + 
  labs(x= "Organisation", y="Hits per Article", title = "The Guardian") + 
  theme_minimal() + 
  ylim(0,4.2)

summary_nyt / summary_guardian +
  plot_annotation(title = "Organisations",
                  theme = theme(plot.title = element_text(size=18)))
```          

## NATO

```{r, message = F, warning = F}
nato <- df_org %>% 
  filter(ORG == "NATO") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "NATO") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2000-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")
  
nato
```

## Taliban
```{r, message = F, warning = F}
taliban <- df_org %>% 
  filter(ORG == "Taliban") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Taliban") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2000-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")
  
taliban
```

## Al Qaeda
```{r, message = F, warning = F}
qaeda <- df_org %>% 
  filter(ORG == "Qaeda") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Al Qaeda") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("2000-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")
  
qaeda
```

## Hamas
```{r, message = F, warning = F}
hamas <- df_org %>% 
  filter(ORG == "Hamas") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Hamas") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1999-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")
  
hamas
```

## Hezbollah
```{r, message = F, warning = F}
hezbollah <- df_org %>% 
  filter(ORG == "Hezbollah") %>% 
  ggplot() +
  aes(x=month, y=penalized_count, color = newspaper) + 
  geom_point(size = 0.9) +
  geom_smooth(alpha=0.2, size = 1.2, method = "gam") +
  labs(y="Hits per Article", x = "Date", title = "Hezbollah") +
  facet_wrap(~newspaper, nrow = 2) +
  scale_x_date(limits = as.Date(c("1999-01-01", "2021-11-01"))) +
  theme_minimal() +
  scale_color_manual(values = color_palette_newspaper, name = "") +
  theme(legend.position="none")
  
hezbollah
```

## Combined Plots for all Key Organisations

### New York Times

All organizations for New York Times.
```{r, message = F, warning = F, fig.dim=c(10,5)}
nato <- df_org %>% 
  filter(ORG == "NATO") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    organisation = "NATO"
  ) 

taliban <- df_org %>% 
  filter(ORG == "Taliban") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    organisation = "Taliban"
  ) 

qaeda <- df_org %>% 
  filter(ORG == "Qaeda") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    organisation = "Al Qaeda"
  ) 

hamas <- df_org %>% 
  filter(ORG == "Hamas") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    organisation = "Hamas"
  ) 

hezbollah <- df_org %>% 
  filter(ORG == "Hezbollah") %>%
  filter(newspaper == "New York Times") %>% 
  mutate(
    organisation = "Hezbollah"
  ) 

pres <- rbind(nato, taliban, qaeda, hamas, hezbollah)

pres <- as_tibble(pres) %>% 
  mutate(organisation = as.factor(organisation))

pres <- pres %>% 
  mutate(organisation = fct_relevel(organisation, c("NATO", "Taliban", "Al Queda", "Hamas")))

pres_nyt <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = organisation) + 
  geom_point(size = 0.5) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Organisations Mentioned in New York Times") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "8 years", date_labels = "%Y") +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "Darjeeling1"), name = "")

pres_nyt
```

### The Guardian

All organizations for the Guardian.
```{r, message = F, warning = F, fig.dim=c(10,5)}
nato <- df_org %>% 
  filter(ORG == "NATO") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    organisation = "NATO"
  ) 

taliban <- df_org %>% 
  filter(ORG == "Taliban") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    organisation = "Taliban"
  ) 

qaeda <- df_org %>% 
  filter(ORG == "Qaeda") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    organisation = "Al Qaeda"
  ) 

hamas <- df_org %>% 
  filter(ORG == "Hamas") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    organisation = "Hamas"
  ) 

hezbollah <- df_org %>% 
  filter(ORG == "Hezbollah") %>%
  filter(newspaper == "The Guardian") %>% 
  mutate(
    organisation = "Hezbollah"
  ) 

pres <- rbind(nato, taliban, qaeda, hamas, hezbollah)

pres <- as_tibble(pres) %>% 
  mutate(organisation = as.factor(organisation))

pres <- pres %>% 
  mutate(organisation = fct_relevel(organisation, c("NATO", "Taliban", "Al Qaeda", "Hamas", "Hezbollah")))

pres_guardian <- pres %>% 
  ggplot() + 
  aes(x=month, y=penalized_count, color = organisation) + 
  geom_point(size = 0.5) +
  geom_smooth(method="loess", se = F, size = 1.2) +
  labs(y="Hits per Article", x = "Date", title = "Organisations Mentioned in The Guardian") +
  scale_x_date(limits = as.Date(c("1996-01-01", "2021-11-01")), breaks = "8 years", date_labels = "%Y") +
  theme_minimal() +
  scale_color_manual(values = wes_palette(name = "Darjeeling1"), name = "")

pres_guardian
```

#### Both Newspapers Without Facet Wrap
```{r, message = F, warning = F}
pres_nyt_y <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,14) 


pres_guardian_y <- pres_guardian +
  theme(legend.position="bottom") +
  ylim(-0.5,14) 


#combining into single plot
pres_both_newspapers <- pres_nyt_y / pres_guardian_y

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers
```

#### Both Newspapers With Facet Wrap
```{r, message = F, warning = F}
pres_nyt <- pres_nyt +
  theme(legend.position = "none") +
  ylim(-0.5,10) +
  facet_wrap(~organisation, nrow = 1) + 
  labs(title = "New York Times")
  
pres_guardian <- pres_guardian + 
  theme(legend.position="none") + 
  ylim(-0.5,10) +
  facet_wrap(~organisation, nrow = 1) + 
  labs(title = "The Guardian")


#combining into single plot
pres_both_newspapers <- pres_nyt / pres_guardian

#removing axis text from the first plot
pres_both_newspapers[[1]] = pres_both_newspapers[[1]] + theme(axis.text.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.title.x = element_blank() )

pres_both_newspapers +
  plot_annotation(title = "Timeline of Organisations",
                  theme = theme(plot.title = element_text(size=18)))
```





