
# Gender Dataset for further analysis

<style>
div.red { background-color:#e06666; border-radius: 5px; padding: 20px;}
</style>
<div class = "red">
This Chapter is not used for the project, but it can be used for further analysis. 
</div>


Here i will make another dataset from `NYT_raw` which also contains the name of the authors and their gender. This dataset called `NYT_clean_author` is *not* used for analysis in this project, but is available for analysis in another project for another time maybe.

It is necessary to make an additional dataset because not all articles has information about the authors. So `NYT_clean_author` will contain less articles than `NYT_clean` because articles with missing author information is dropped from the dataset. 

## Dataset with authors

First thing first, we start by loading the packages `tidyverse` [@tidyverse2019], `DT` [@R-DT] and `gender` [@R-gender].
```{r}
pacman::p_load(tidyverse, gender, DT)
```

We are going to make the author dataset `NYT_clean_author` from the dataset `NYT_raw`. It mostly follows the syntax from the package `tidyverse`.
```{r cleaning the data with author, eval = F}
#selecting coloumns to keep
coloumns_to_select <- c("response.docs.headline.main", "response.docs.web_url",
"response.docs.pub_date", "firstname", "lastname", "rank",)

#creating the new dataframe
NYT_aut <- NYT_raw %>% 
  # Information about the name of the author is stored inside a dataframe in a column of the dataframe all_articles. I use the function "unnest" to unpack the dataframe into separate columns
  unnest(cols = response.docs.byline.person) %>% 
  
  #selecting the defined columns
  select(coloumns_to_select) %>% 
  
  #renaming columns to more humane names
  rename(
    "headline" = "response.docs.headline.main",
    "url" = "response.docs.web_url",
    "date" = "response.docs.pub_date"
  ) %>% 
  
  #making a new coloumn with full name
  mutate(
    full_name = str_c(firstname, lastname, sep = "_")
  ) %>% 
  
  #formating coloumns to the correct class
  mutate(
    date = as.Date(date),
    firstname = as.factor(firstname),
    lastname = as.factor(lastname),
    rank = as.factor(rank),
    full_name = as.factor(full_name)
  ) %>% 
  
  #filtering rows where author is missing
  filter(is.na(full_name) == F) %>% 
  
  #arranging by date so that the articles are in chronological order
  arrange(by=date)


#saving to a csv
write_csv(NYT_aut, "data/new_york_times/data_additional/NYT_clean_author_cp18.csv")
```


Now lets inspect the cleaned dataframe. to see if it looks allright. Again, we use the function `datatable` from the package `DT`. The cleaned dataset contains the following columns. 

* `headline` which is the title/headline of the article
* `url` which is a url leading to the the article on the NYT webpage
* `date` which is the publication date
* `firstname` which is the first name of the author of the article
* `lastname` which is the last name of the author of the article
* `rank` which indicates the rank of the author.
* `full_name` which is the full name of the author

```{r inspecting the cleaned dataframe, message = F, warning = F}
#reading the csv
NYT_aut <- read_csv("data/author_dataset/NYT_author.csv")

#making a nice dataframe that we can browse. Note that we remove abstract because there is too much text in it to show in a nice way.
font.size <- "8pt"
   DT::datatable(
     NYT_aut,
     rownames = FALSE,
     filter = "top", 
     options = list(
       initComplete = htmlwidgets::JS(
          "function(settings, json) {",
          paste0("$(this.api().table().container()).css({'font-size': '", font.size, "'});"),
          "}"),
       pagelength = 3,
       scrollX=T,
       autoWidth = TRUE
       ) 
     )

```

This dataset `NYT_aut` has another structure than `NYT_clean`. In `NYT_aut` each row corresponds to an author instead of an article. This means that an article has *x* number of rows for *x* number of authors to that article. The column `rank` indicates the rank of each author, so that an article with 2 authors will have 2 rows with 1 and 2 as values in rank. This is because the function `unnest` makes x number of observations for x number of authors. You can see it for yourself in the table below. 
```{r, warning = F}
#dropping abstract and snippet since they make it hard to read the data
NYT_aut %>%
  #filtering rows where rank=1 is followed by rank=2 and where rank > 1
  filter(lead(rank) == 2 | rank > 1) %>% 
  #arranging by date
  arrange(url) %>% 
  #making a datatable again
   DT::datatable(
     rownames = FALSE,
     filter = "top", 
     options = list(
       initComplete = htmlwidgets::JS(
          "function(settings, json) {",
          paste0("$(this.api().table().container()).css({'font-size': '", font.size, "'});"),
          "}"),
       pagelength = 3,
       scrollX=T,
       autoWidth = TRUE
       ) 
     )
```


## Comparing the datasets 
Lastly, we are going to do a quick comparison of the datasets `NYT_clean` and `NYT_aut`. We start by loading ``
```{r}
df_NYT <- read_csv("data/new_york_times/data_additional/NYT_clean_cp2.csv") 
```

We start by comparing the number of articles in each dataset.
```{r}
#summing articles in df_nyt
nrow(df_NYT)

#summing articles in NYT_aut
NYT_aut %>% 
  filter(rank == 1) %>% 
  nrow()

```
We see that `df_NYT` has 23.502 articles similar to `data_raw` and `NYT_aut` has 18.417 articles meaning that ~7000 articles has been dropped in the author dataset. This is the reason why we work with two datasets instead of one.

In the next chunk we see of there are specific dates where articles from `NYT_aut` has been dropped. In other words if there is an unbalance in the dates of the author dataset. 
```{r comparingdatasets, message = F, warning = F, figwidth = 12, fig.cap = "Number of articles published each day in the full dataset and in the dataset with authors."}
#defining a dataset from df_NYT containing only dates and a indicator
x1 <- df_NYT %>% 
  select(date) %>% 
  mutate(
    Dataset = "Full Dataset"
  )

#defining a dataset from NYT_aut containing only dates and a indicator
x2 <- NYT_aut %>% 
  filter(rank == 1) %>% 
  select(date) %>% 
  mutate(
    Dataset = "Dataset with authors"
  )

#contatinating the two datasets and setting data to factor
xx <- as_tibble(rbind(x1, x2)) %>% 
  mutate(
    Dataset = as.factor(Dataset)
  )

#relevelling Dataset
xx$Dataset <- relevel(xx$Dataset, "Full Dataset")

#we sum together articles for each date and make a plot for each dataset
xx %>%
  group_by(date, Dataset) %>%
  tally() %>% 
  ggplot() +
  aes(x=date, y=n, color = Dataset) +
  geom_line() +
  labs(y = "hits")

```
By eyeballing figure \@ref(fig:comparingdatasets) we see that `NYT_aut` is missing articles at a fairly evenly rate across all dates. In other words there are not specific periods where `NYT_aut` drops completely in articles. The ~7000 articles that are missing from `df_NYT` are spread across the whole period.

```{r, include = F, eval = F}
#relevelling Dataset
xx$Dataset <- relevel(xx$Dataset, "Dataset with authors")

#another plot sohwing the difference...
xx %>%
  group_by(date, Dataset) %>%
  tally() %>% 
  ggplot() +
  aes(x=date, y=(n-lead(n))) +
  geom_line() +
  labs(y = "hits")
```



## Adding gender information to `NYT_aut`
Next up we are going to add a gender to each author (M/F). I found a package 
named `gender` which contains lists of male and female names. We can compare the
names in this dictionary to the name of the authors in our dataset and assign
each author a gender. First we reformat the dataframe a little bit. 
```{r, eval = F}
#reformatting the dataframe so it can enter the "gender_df" function below
NYT_aut <- NYT_aut %>%  mutate(
  firstname = as.character(firstname),
  year = 2012
)
```

Then we use the function `gender_df` to predict the gender of all the authors in our dataset. 
```{r, warning = F, eval= F}
#the function gender_df predicts gender using a coloumn of names from a dataframe
gender_basic <- gender_df(
  NYT_aut,
  name_col = "firstname",
  year_col = "year"
)
```


Then we merge the dataframes `gender_basic` and `NYT_aut`. This is a little messy but it works. 
```{r, eval = F}
#renaming the column containing names so it can merge back with df
gender_basic <- gender_basic %>%
  rename(firstname = name)

#joining the two dataframes "NYT_aut" and "gender_basic". NYT_aut contains duplicates of
#author, since each author wrote several articles in the period where the data
#was collected. We deal with these duplicates by adding a row number, so
#each author gets two keys to join by. 
NYT_aut <- left_join(NYT_aut %>% group_by(firstname) %>% mutate(id = row_number()),
          gender_basic %>% group_by(firstname) %>% mutate(id = row_number()), 
          by = c("firstname", "id"))
```

Finally we remove the extra columns created by the function `gender_df` which we dont need. 
```{r, message = F, eval = F}
#removing the extra columns created by the function "gender_df"
coloumns_to_remove = c("year", "id", "proportion_male", "proportion_female", "year_min", "year_max")

NYT_aut <- NYT_aut %>% 
  select(-(coloumns_to_remove))
```

We save the dataframe
```{r, eval = F}
write_csv(NYT_aut, "data/new_york_times/data_additional/NYT_clean_author_cp18.csv")
```


Now we have successfully assigned a gender to each author. Lets inspect this. Let's see how many male, female and missing authors there are in the dataset.      
```{r gendertable}
#checking how many males, females and NA's there are
NYT_aut %>%
  #removing duplicates
  distinct(full_name, .keep_all = TRUE) %>% 
  group_by(gender) %>%
  summarize(count = n()) %>% 
  knitr::kable(caption = "The number of authors by gender",
               col.names = c("Gender", "Count"))
```

Table \@ref(tab:gendertable) shows that there are 764 female authors and 1610 male authors, meaning that there are approximately 2 times male authors compared to female authors. This is quite the difference. 

Moreoever, there are 370 NA's meaning that the names of 370 authors could not be found in the database from the package `gender`. What is the deal with these NA's? Lets see which author names cannot be assigned to a gender (M/F). 
```{r}
NYT_aut %>%
  distinct(full_name, .keep_all = TRUE) %>% 
  filter(is.na(gender)) %>% 
  select(firstname, full_name) %>% 
  head(15) %>% 
  knitr::kable(caption = "The firstname and full names of authors where the gender-package failed to assign a gender",
               col.names = c("First name", "Full Name"))
  
```
Here we can see the firstname of authors which the gender-package failed to
recognize. These seem like the names of immigrants and some weird names. 




Some more gender stuff over time
```{r}
NYT_aut %>% 
  group_by(date, gender) %>% 
  tally() %>%
  ggplot() + 
  aes(x=date, y=n, color = gender) + 
  geom_line() +
  labs(y = "hits") + 
  facet_wrap(~gender, nrow = 3)

```




